---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dquad

This is a support repository for the paper *Bias Correction of Quadratic Spectral Estimators*. The code is mainly implemented in the <tt>speccy</tt> package, found [here](https://github.com/astfalckl/speccy). This package is in active development to house many modern developments in spectral analsis, including this body of work. This document replicates the examples in Figure~1 of the paper to demonstrate implmentation of the methodology.

# Generate data

```{r message = FALSE}
# Needed for calculation
library(speccy)

# Needed for this example to obtain the dpss tapers
library(multitaper)

# Handy for data manipulation and plotting
library(tidyverse)
library(latex2exp)
library(patchwork)
```

```{r echo = FALSE}
set.seed(2)
theme_set(theme_bw())
```

```{r}
overlap <- 0
m <- 2^5
l <- 2^9
k <- l / 4

s <- l * (1 - overlap)
n <- welch_length(m, l, p = overlap)

k <- l / 4

phis <- c(2.7607, -3.8106, 2.6535, -0.9238)
sd <- 1
delta <- 1

sampled_ar <- stats::arima.sim(
    list(ar = phis), n, n.start = 1000, sd = sd
)
```

# Calculate standard spectral estimates

As we detail in the paper, the three broad families of spectral estimator are multitaper, lag-window and Welch estimates. We demonstrate the bias correction methodology to each of these estimators, calculated over the same time-series. First we parameterise the estimators. The choices of tapers/windows were made so that each estimator exhibited similar bias.

```{r}
# Welch
h <- gsignal::hamming(l)

welch_estimate <- welch(sampled_ar, m, l, s, delta = delta, h = h)

welch_biased_spec <- bochner(
    ar_acf(l, phis, sd), h = h,
    delta = delta, return_ff = FALSE
)

welch_fit <- tibble(
    ff = welch_estimate$ff,
    estimate = welch_estimate$estimate,
    true_spec = ar_spectrum(welch_estimate$ff, phis, sd, delta),
    biased_spec = welch_biased_spec
)

#  Multitaper set-up
slepian_tapers <- dpss(n, 2 * m, m)$v

mt_estimate <- multitaper(sampled_ar, slepian_tapers, delta = delta)

mt_expectation <- multitaper(
  ar_acf(n, phis, sd), slepian_tapers, acf = TRUE, delta = delta
)

multitaper_fit <- tibble(
    ff = mt_estimate$ff,
    estimate = mt_estimate$estimate,
    true_spec = ar_spectrum(mt_estimate$ff, phis, sd, delta),
    biased_spec = mt_expectation$estimate
)

# Lag window set-up
w <- w_daniell(n, 2 * l)

lw_estimate <- lag_window(sampled_ar, delta = delta, lag_sequence = w)

lw_biased_spec <- lag_window(
  ar_acf(n, phis, sd),
  delta = delta, lag_sequence = w, return_ff = FALSE, acf = TRUE
)

lag_window_fit <- tibble(
    ff = lw_estimate$ff,
    estimate = lw_estimate$estimate,
    true_spec = ar_spectrum(lw_estimate$ff, phis, sd, delta),
    biased_spec = lw_biased_spec
)
```

# Calculate bias corrected estimates

Calculating the bias corrected estimates requires two steps. First build the bases. These are a function only of the parameterisation of the estimators (time-length, taper/window selection etc). In application, the bases can be pre-computed for all estimators of a given parameterisation. The second step is calculating the solution to the weighted least squares problem, as described in the paper.

```{r, message = FALSE, warning = FALSE, results = "hide"}
B_welch <- speccy::build_bases(l, k = k, h = h, diff_order = 1)
B_multi <- speccy::build_bases(n, k = k, h = slepian_tapers, diff_order = 1)
B_lag <- speccy::build_bases(n, k = k, lag_sequence = w, diff_order = 1)

dwelch_estimate <- speccy::debias_quad(welch_fit$estimate, B_welch)
dmulti_estimate <- speccy::debias_quad(multitaper_fit$estimate, B_multi)
dlag_estimate <- speccy::debias_quad(lag_window_fit$estimate, B_lag)

dwelch_estimate <- tibble(
  estimate = dwelch_estimate$estimate,
  ff = dwelch_estimate$ff
)

dmulti_estimate <- tibble(
  estimate = dmulti_estimate$estimate,
  ff = dmulti_estimate$ff
)

dlag_estimate <- tibble(
  estimate = dlag_estimate$estimate,
  ff = dlag_estimate$ff
)
```

# Plot results

Finally, we can plot the results. Here, I have used <tt>ggplot2</tt> which has a habit of being very verbose and so I've suppressed the code in the README. Please consult the raw markdown file if you'd like to access any of the plotting functions.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, fig.height = 3}
inset_y_min <- 2.5e-3
inset_y_max <- 1.5e-1

welch_inset <- ggplot(welch_fit) +
  geom_line(
    aes(x = ff, y = estimate),
    colour = "darkgrey", linewidth = 0.5, alpha = 0.5
  ) +
  geom_line(
    data = dwelch_estimate,
    aes(x = ff, y = estimate), linewidth = 0.5, alpha = 0.5
  ) +
  geom_line(
    aes(x = ff, y = true_spec),
    linetype = "dashed", linewidth = 0.5
  ) +
  geom_line(
    aes(x = ff, y = biased_spec),
    linetype = "dashed", linewidth = 0.5, colour = "darkgrey"
  ) +
  scale_x_continuous(
    limits = c(0.3, .5),
    breaks = c(0.375, 0.5),
    labels = c(TeX("$3\\pi/4$"), TeX("$\\pi$"))
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    trans = "log10",
    limits = c(inset_y_min, inset_y_max),
    breaks = c(1e-1, 1e-2),
    labels = c(TeX("$10^{-1}$"), TeX("$10^{-2}$"))
  ) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank()
  )

welch_plot <- ggplot(welch_fit) +
  annotation_custom(
    ggplotGrob(welch_inset),
    xmin = 0.2, xmax = 0.5,
    ymin = log10(4 * 10^{-1}), ymax = log10(3 * 10^5)
  ) +
  annotate(
    "rect",
    xmin = 0.3, xmax = 0.5, ymin = inset_y_min, ymax = inset_y_max,
    alpha = 0.1, fill = "black", colour = "black", linetype = "dotted"
  ) +
  geom_line(
    aes(x = ff, y = estimate),
    colour = "darkgrey", linewidth = 0.5, alpha = 0.5
  ) +
  geom_line(
    data = dwelch_estimate,
    aes(x = ff, y = estimate), linewidth = 0.5, alpha = 0.5
  ) +
  geom_line(
    aes(x = ff, y = true_spec),
    linetype = "dashed", linewidth = 0.5
  ) +
  geom_line(
    aes(x = ff, y = biased_spec),
    linetype = "dashed", linewidth = 0.5, colour = "darkgrey"
  ) +
  scale_y_continuous(
    TeX("Spectral density [$m^2 s]"),
    trans = "log10",
    limits = c(0.1 * 10^-2, 10 * 10^4),
    breaks = c(10^-2, 10^0, 10^2, 10^4),
    labels = c(
      TeX("$10^{-2}$"),
      TeX("$10^0$"),
      TeX("$10^2$"),
      TeX("$10^4$")
    )) +
  scale_x_continuous(
    "Frequency [rad/s]",
    limits = c(0, 0.5),
    expand = c(0,0),
    breaks = c(0, 0.125, 0.25, 0.375, 0.5),
    labels = c(
      TeX("$0$"),
      TeX("$\\pi/4$"),
      TeX("$\\pi/2$"),
      TeX("$3\\pi/4$"),
      TeX("$\\pi$")
    )) +
  ggtitle("Welch Estimate") +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    )

multitaper_inset <- multitaper_fit %>%
  slice(which(row_number() %% m == 1)) %>%
  ggplot() +
    geom_line(
      aes(x = ff, y = estimate),
      colour = "darkgrey", linewidth = 0.5, alpha = 0.5
    ) +
    geom_line(
      data = dmulti_estimate,
      aes(x = ff, y = estimate), linewidth = 0.5, alpha = 0.5
    ) +
    geom_line(
      aes(x = ff, y = true_spec), linewidth = 0.5, linetype = "dashed"
    ) +
    geom_line(
      aes(x = ff, y = biased_spec), linewidth = 0.5, linetype = "dashed",
      colour = "darkgrey"
    ) +
    scale_x_continuous(
      limits = c(0.3, .5),
      breaks = c(0.375, 0.5),
      labels = c(TeX("$3\\pi/4$"), TeX("$\\pi$"))
    ) +
    scale_y_continuous(
      expand = c(0, 0),
      trans = "log10",
      limits = c(inset_y_min, inset_y_max),
      breaks = c(1e-1, 1e-2),
      labels = c(TeX("$10^{-1}$"), TeX("$10^{-2}$"))
    ) +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank()
    )

multitaper_plot <- multitaper_fit %>%
  slice(which(row_number() %% m == 1)) %>%
  ggplot() +
    annotation_custom(
      ggplotGrob(multitaper_inset),
      xmin = 0.2, xmax = 0.5,
      ymin = log10(4 * 10^{-1}), ymax = log10(3 * 10^5)
    ) +
    annotate(
      "rect",
      xmin = 0.3, xmax = 0.5, ymin = inset_y_min, ymax = inset_y_max,
      alpha = 0.1, fill = "black", colour = "black", linetype = "dotted"
    ) +
    geom_line(
      aes(x = ff, y = estimate),
      colour = "darkgrey", linewidth = 0.5, alpha = 0.5
    ) +
    geom_line(
      data = dmulti_estimate,
      aes(x = ff, y = estimate), linewidth = 0.5, alpha = 0.5
    ) +
    geom_line(
      aes(x = ff, y = true_spec), linewidth = 0.5, linetype = "dashed"
    ) +
    geom_line(
      aes(x = ff, y = biased_spec), linewidth = 0.5, linetype = "dashed",
      colour = "darkgrey"
    ) +
    scale_y_continuous(
      TeX("Spectral density [$m^2 s]"),
      trans = "log10",
      limits = c(0.1 * 10^-2, 10 * 10^4),
      breaks = c(10^-2, 10^0, 10^2, 10^4),
      labels = c(
        TeX("$10^{-2}$"),
        TeX("$10^0$"),
        TeX("$10^2$"),
        TeX("$10^4$")
      )) +
    scale_x_continuous(
      "Frequency [rad/s]",
      limits = c(0, 0.5),
      expand = c(0,0),
      breaks = c(0, 0.125, 0.25, 0.375, 0.5),
      labels = c(
        TeX("$0$"),
        TeX("$\\pi/4$"),
        TeX("$\\pi/2$"),
        TeX("$3\\pi/4$"),
        TeX("$\\pi$")
      )) +
    ggtitle("Multitaper Estimate") +
      theme_bw() +
      theme(
        panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
      )

lag_inset <- lag_window_fit %>%
  slice(which(row_number() %% m == 1)) %>%
  ggplot() +
      geom_line(
      aes(x = ff, y = estimate),
      colour = "darkgrey", linewidth = 0.5, alpha = 0.5
    ) +
    geom_line(
      data = dlag_estimate,
      aes(x = ff, y = estimate), linewidth = 0.5, alpha = 0.5
    ) +
    geom_line(
      aes(x = ff, y = true_spec), linewidth = 0.5, linetype = "dashed"
    ) +
    geom_line(
      aes(x = ff, y = biased_spec), linewidth = 0.5, linetype = "dashed",
      colour = "darkgrey"
    )  +
    scale_x_continuous(
      limits = c(0.3, .5),
      expand = c(0, 0),
      breaks = c(0.375, 0.5)
    ) +
    scale_y_continuous(
      expand = c(0, 0),
      trans = "log10",
      limits = c(inset_y_min, inset_y_max),
      breaks = c(1e-2)
    ) +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank()
    )

lag_plot <- lag_window_fit %>%
  slice(which(row_number() %% m == 1)) %>%
  ggplot() +
  annotation_custom(
    ggplotGrob(lag_inset),
    xmin = 0.2, xmax = 0.5,
    ymin = log10(4 * 10^{-1}), ymax = log10(3 * 10^5)
  ) +
  annotate(
    "rect",
    xmin = 0.3, xmax = 0.5, ymin = inset_y_min, ymax = inset_y_max,
    alpha = 0.1, fill = "black", colour = "black", linetype = "dotted"
  ) +
  geom_line(
    aes(x = ff, y = estimate),
    colour = "darkgrey", linewidth = 0.5, alpha = 0.5
  ) +
  geom_line(
    data = dlag_estimate,
    aes(x = ff, y = estimate), linewidth = 0.5, alpha = 0.5
  ) +
  geom_line(
    aes(x = ff, y = true_spec), linewidth = 0.5, linetype = "dashed"
  ) +
  geom_line(
    aes(x = ff, y = biased_spec), linewidth = 0.5, linetype = "dashed",
    colour = "darkgrey"
  ) +
  scale_y_continuous(
    TeX("Spectral density [$m^2 s]"),
    trans = "log10",
    limits = c(0.1 * 10^-2, 10 * 10^4),
    breaks = c(10^-2, 10^0, 10^2, 10^4),
    labels = c(
      TeX("$10^{-2}$"),
      TeX("$10^0$"),
      TeX("$10^2$"),
      TeX("$10^4$")
    )) +
  scale_x_continuous(
    "Frequency [rad/s]",
    limits = c(0, 0.5),
    expand = c(0,0),
    breaks = c(0, 0.125, 0.25, 0.375, 0.5),
    labels = c(
      TeX("$0$"),
      TeX("$\\pi/4$"),
      TeX("$\\pi/2$"),
      TeX("$3\\pi/4$"),
      TeX("$\\pi$")
    )) +
  ggtitle("Lag-window Estimate") +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      axis.title.x = element_blank()
    )

plot_final <- lag_plot + multitaper_plot + welch_plot
# plot_final <- lag_plot + welch_plot

plot_final
```